Dynamics of the Lower Solar Atmosphere

### 2024-08-14 chapter6/ms.tex
Updates from Overleaf

%\begin{abstract}

 %   The \ion{Mg}{2}~k \& h line intensity ratios can be used to probe the characteristics of the plasma in the solar atmosphere. In this study, using the observations recorded by the Interface Region Imaging Spectrometer (IRIS), we study the variation of the \ion{Mg}{2}~k \& h intensity ratio for three flares belonging to X-class, M-class, and C-class, throughout their evolution. We also study the k-to-h intensity ratio as a function of magnetic flux density obtained from the line-of-sight magnetograms recorded by the Helioseismic and Magnetic Imager (HMI) on board the Solar Dynamics Observatory (SDO). Our results reveal that while the intensity ratios are independent of magnetic flux density, they show significant changes during the evolution of the C-class and M-class flares. The intensity ratios start to increase at the start of the flare and peak during the impulsive phase before the flare peak and decrease rapidly thereafter. The values of the ratios fall even below the pre-flare level during the peak and decline phases of the flare. These results are important in the light of heating and cooling of localized plasma and provide further constraint on the understanding of flare physics.

%\end{abstract}

%%----------------------------------------------------
\section{Introduction} \label{sec:intro}
%%----------------------------------------------------

Solar flares are the most energetic events on the Sun, where an enormous amount of magnetic free energy is released due to the reconfiguration of the coronal magnetic field. The released energy can cause particle acceleration, heating and flows in the solar atmosphere and a transient enhancement in solar radiative output. It is observed that most of the energy radiated in flares originates from the dense chromosphere \citep{fletcher10,milligan14}. Hence, studying the chromospheric lines during flares provides us with diagnostics, which may be important for understanding the physics of solar flares and their effect on the local plasma environment.

The chromosphere emits in various UltraViolet(UV) and optical lines. While many optical lines, e.g., H$\alpha$, \ion{Ca}{2}, are routinely observed from ground-based telescopes, observations in the \ion{Mg}{2} resonance lines have been rare in the past. Since the launch of the Interface Region Imaging Spectrograph \citep[IRIS;][]{IRIS} we have been in the position to monitor these lines regularly with excellent spatial and spectral resolution.

The \ion{Mg}{2}~k and h lines are transitions to the ground state from a finely split pair of upper levels ($3p~^{2}P_{\nicefrac{3}{2}}${--}$3s~^{2}S_{\nicefrac{1}{2}}$ and $3p ^{2}P_{\nicefrac{1}{2}}${--}$3s~^{2}S_{\nicefrac{1}{2}}$). These transitions create the optically thick lines in the wavelengths 2796.34~{\AA} (\ion{Mg}{2}~k) and 2803.52~{\AA} (\ion{Mg}{2}~h). It is suggested that the intensity ratios of these lines can be used to probe the optical depth of the local environment \citep{kerr15}.

The integrated intensity of a line transitioning from an upper-level j to a lower-level i, is dependent on the collision strength for that transition $\Omega_{ij}$, which is given by~\citep{henri62,mariska92},

%%---------------------------------------------------------------
\begin{equation*}
    \Omega_{ij}~=~\frac{8\pi}{\sqrt{3}}~\frac{I_{H}}{\Delta \epsilon_{ij}}~g~\omega_{i}~f_{ij}
\end{equation*}
%%---------------------------------------------------------------

\noindent Where $I_{H}$ is the ionization energy of Hydrogen, $\Delta \epsilon_{ij}$ is the threshold energy for the transition, g is the Gaunt factor, $\omega_{i}$ is the statistical weight of the level and $f_{ij}$ is the oscillator strength. In optically thin conditions, the intensity ratio of the k to h line is the ratio of the collision strengths, as the escape probability of photon is unity. The \ion{Mg}{2} k and h lines share the same ionization state and originate from a transition to a shared lowered level. As the statistical weight ($\omega_{i}$) is same in both cases, the line intensity ratio is simply the ratio of the oscillator strengths($f_{ij}$). This implies the ratio is 2:1 in optically thin conditions, and lower when the medium is optically thick \citep{kerr15,levens19}.

In addition, the \ion{Mg}{2} k and h lines can be used to estimate the velocity in the middle and upper chromosphere, the chromospheric velocity gradients, the temperature in the middle chromosphere \citep{leenarts13a,leenarts13b,pereira13}. The \ion{Mg}{2} triplets in emission can be used to identify heating in the lower chromosphere \citep{pereira15}. There have also been multiple studies that have shown a spatial variation of the \ion{Mg}{2} line profiles \citep{dalda23,panos18}. \cite{polito23} showed that the leading edge of the flare ribbon is associated with enhanced broadening and strong central reversal. They interpreted the difference in the profile as a difference in the heating mechanism at the leading edge and bright part of the flare ribbons. \cite{panos21,panos21_2} showed similar differences between the line profiles and energy input.

Using the observations recorded by the OSO-8 LPSP instrument, \citep{lemaire84} studied the evolution of the intensity ratio of \ion{Mg}{2}~h~\&~k, \ion{Ca}{2}~h~\&~k and Ly~$\alpha$~\&~$\beta$ lines. \cite{lemaire84} showed that the intensity ratio of the \ion{Ca}{2}~k/h lines increased from 1 to 1.2 during the ascending phase of a flare and returned to 1 during the later phases. The authors interpreted the correlated temporal behavior across various elements as an indication of downward energy propagation. We note that this may suggest a slight decrease in the opacity due to localized heating at the formation height of the \ion{Ca}{2} line during the rise phase of the flare.

%\sout{In spite of the excellent potential of these spectral lines in diagnosing the localized atmosphere, such} \sout{studies have not been taken up in detail due to the scarcity of spectral observations in these lines.}

Here, we study the evolution of the intensities ratios of the \ion{Mg}{2}~h~\&~k lines during the course of the evolution of three flares, \textit{viz.}, C-class, M-Class and X-class. We focus on the the dependence of line ratios on the underlying magnetic field strength, which to our knowledge has not been explored so far. The rest of the paper is structured as follows. \S\ref{sec:obs} discusses the observations used in this paper. We discuss how we reduce and analyze the data and the results in \S\ref{sec:dar}.

>>>>

\justifying

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Solar flares are the largest magnetic events in our solar system. They are characterized by the eruptive release of magnetic energy, and manifest as localized transient heating of coronal plasma to temperatures over $\sim$ 10~MK. Flares are characterized by the full-disk integrated peak soft X-ray flux observed by the X-Ray Sensor onboard the {\it Geostationary Operational Environmental Satellites} \citep[{\it GOES}/XRS,][]{xrs} (``\textit{GOES class}''). The various components of the flare manifest in different wavelengths of the electromagnetic spectrum. The flare arcade and the fan exhibit the thermal plasma, while the majority of the non-thermal energy is from the foot points \citep{benz17}. The hot coronal plasma exhibited by flares is generally attributed to chromospheric evaporation, heated by the collisions of the flare-accelerated electrons in the impulsive phase \citep{fletcher11}. There is also strong evidence that the plasma might be directly heated in the corona in some cases \citep[e.g.][]{longcope11, reeves17}.

There have been several works \citep{stosire07,emslie12,inglis14,warmuth16a,warmuth16b,ash17} that study the partition between the thermal and non-thermal energies in flares, which holds key insight into the relevant underlying physical processes in the flares. The results have been contradictory, and there is no consensus on whether non-thermal energetic electrons have enough energy to power the observed thermal energy component of smaller flares. These studies usually use peak thermal energy as a representative of the bulk thermal output of the flares. While these quantities are fair representation of the bulk thermal output of the flare, identifying subtle differences in the heating mechanisms at play at different stages of the flares requires a quantification of the thermal energy as a function of time.

There are several studies \citep{hilarie05,caspi10} that estimate the thermal energy as a function of time. The thermal energy of the flares is defined as 
\begin{equation}
\mathrm{U_{Th}~\simeq~3n_{e}k_{B}TV}f,
\end{equation} \label{eq:t_eneg_1}
 where `$\mathrm{n_{e}}$' is the electron number density of the flaring plasma, `V' is the volume of the flaring plasma, `\textit{f}' is the volume filling factor, and `T' is the instantaneous temperature. One of the key challenges in characterizing the flares' thermal energy is determining the volume.

\cite{li23} used $\frac{1}{e^{2}}$ contour of soft X-ray emitting region to identify the flare area `A', which was subsequently used to estimate the volume, $V\sim A^{\frac{3}{2}}$. \cite{zhang19} used a similar method, using intensity threshold on AIA 131~{\AA} observation, to estimate the flare area (A). \cite{hilarie05} used a similar method with {\it Reuven Ramaty High Energy Solar Spectroscopic Imager} \citep[{\it RHESSI},][]{rhessi} soft X-ray images to estimate the lower limit of the volume. They used {\it RHESSI} hard X-ray imaging to estimate the distance between two footpoints. Assuming a perfect ``arc-shaped" loop they calculated the upper estimate of the volume and concluded that the largest source of uncertainty in determining the thermal energy arises from the uncertainty in the filling factors.

\cite{hilarie05} also estimated the evolution of cumulative non-thermal energy injected by the electrons by assuming a thick-target model \citep{brown71} and fitting the {\it RHESSI} spectrum of the events. They fitted the {\it RHESSI} spectra in 25{--}35 keV bands to determine the non-thermal energies in time bins with significant emission above 25 keV. This process ensures a better accuracy for determining non-thermal energy by dividing the flare into short time intervals with large energy inputs, resulting in small thermal energy losses. They found that the non-thermal cumulative energy was larger than the thermal component of the flare for most of the duration of the flare. This is expected as radiative cooling significantly lower the temperature of the flare arcade as the flare evolves. They also deduced that during the early impulsive phase, the cumulative non-thermal energy could be lower than the thermal energy, depending on the estimates used for the flaring plasma volume.

The filling factor `\textit{f}' quantifies the fraction of the apparent volume that is filled with the plasma of a specific temperature. It might also hold information about the geometric projection effects of the real volume into the apparent projected volume that is observed. Due to the highly multi-thermal nature of the flaring plasma and rapidly changing volume during the rise phase of the flare, it is challenging to define a filling factor that describes the plasma volume across the evolution of the flare. %Furthermore, most of the statistical studies so far, deals with bolometric or time averaged quantities near the peak of the flare.}
%This would indicate that direct heating possibly happens before or during the impulsive phase of flares.

\cite{caspi10} studied super hot flares from {\it RHESSI} observations, and used a 50\% intensity contour of cleaned {\it RHESSI} images to estimate the flare area, A. The volume $V=\frac{4}{3}\pi (A/\pi)^{3/2}$ was inferred from the area assuming spherical symmetry.  The two temperature components from the {\it RHESSI} spectral fit were assigned a volume V/2 each to estimate the total thermal energy. In all of these cases, the assumption of symmetrical expansion is useful because the thermal energy was estimated from a spectra obtained by integrating multiple pixels. There is no reliable way to address the spatially varying multi-thermal nature of the flaring plasma in these cases.

In addition to the bulk thermal plasma in the flare loop, there are studies that suggested completely different heating mechanisms at play in specific flares. For example, \citet{longcope11} suggested a super hot plasma component in flares is directly heated in the corona in some large solar flares. Sustained temperatures in supra-arcade fans in flares have been hypothesized to be caused by supra-arcade down flows \citep[e.g.][]{reeves17}, suppression of conduction by turbulence \citep[e.g.][]{xie23} or global compression from reconnection inflows \citep[e.g.][]{reeves19}. These processes are distinctly different from chromospheric evaporation, which is characterized by the development of the hard X-ray foot-points and increasing intensity of soft X-ray and EUV-emitting plasma as the loops fill up.

%%%%######%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={0cm 0.5cm 1.2cm 0.5cm},clip,width=\textwidth]{flare_pos.pdf}
    \caption{Position of various instruments during the flares. AIA 193~{\AA} observations during 2021 Oct 28 (panel a) and EUVI 171~{\AA} observation during 2020 Nov 29 (panel c) in Heliographic Stonyhurst (HGS) coordinates. The two flares studied here are enclosed by white dotted circle. The projected positions of {\it Solar Orbiter}, {\it SDO} and {\it STEREO-A} are marked as labelled. Positions of various spacecrafts projected onto the Heliocentric Inertial (HCI) frame for the for 2021 Oct 28 (panel c) and 2020 Nov 29 (panel d), as labelled.}
    \label{fig:sc_pos}
\end{figure}
%%%%######%%%%%%

Here we study the evolution of the thermal energy of two flares, \textit{viz.} M-class and X-class. We use existing tools and observations of the same region on the Sun from different vantages to estimate the volume of the flaring plasma more accurately. We show that a better estimate of the plasma volume affects the estimate of the thermal energy and has implications towards the energy budget of the flare. We also estimate the cumulative non-thermal energy deposited by the non-thermal electrons and compare it with our thermal energy estimates for the flare observed on 28 October, 2021. In \S\ref{sec:obs} we discuss the observations used in this paper and details of our analysis. We discuss our method for estimating the thermal energy in \S\ref{sec:therm}. In \S\ref{sec:los} and \S\ref{sec:non-therm}, we describe our method of volume estimation and cumulative non-thermal energy estimation, respectively. We discuss the results in \S\ref{res}.

%%%%%%
\begin{table}
    \centering
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{||c|c|c||}
    \hline
    Instruments & Details & Flares\\
    \hline 
    \hline
      & 0.6 \arcsec/pix. Used&   \\
     {\it SDO}/AIA &  for estimating the thermal & 28 October, 2021 \\
     & properties. Temperature coverage & \& 29 November, 2020 \\
     & 5 $< log\ (T) <$ 7.5 & \\
         \hline
        & 2.5 \arcsec/pix. Has & \\
       & similar temp. response as AIA. & 28 October 2021 \\
      {\it GOES}/SUVI & Used for replacing saturated & \& 29 November, 2020 \\
       & AIA frames in DEM estimation & \\
       \hline 
        & 1 \arcsec/pix. Used simultaneously & \\
       & with AIA data in estimation of the thermal  & \\
     {\it Hindoe}/XRT  & properties. Temperature coverage  & 29 November, 2020 \\
       & 6.1 $< log\ (T) <$ 7.5, provides better & \\
       & constraints on the higher temp. end. & \\
       \hline 
       & 1.6 \arcsec/pix. Used & \\
       & for estimating the geometry & 28 October 2021  \\
     {\it STEREO-A}/EUVI   & of the arcade. 171 and 195 \AA & \& 29 November, 2020 \\
       & Observations are used with & \\
       & AIA 171 and 193 \AA. & \\
       \hline
        & X-ray imaging and spectroscopy in the  &   \\
       & 4-150 keV range. Soft X-ray imaging used & \\
      {\it SO}/STIX & for calculating the area of the flare. & 28 October 2021 \\
       &  Spectra used for constraining the thermal & \\
       & and non-thermal properties & \\
       \hline
       & Continuous full disk integrated soft X-ray light curve. & 28 October, 2021 \\
     {\it GOES}/XRS  & Used for estimating the bulk thermal & \& 29 November, 2020 \\
       & properties of the flare. & \\
       \hline
    \end{tabular}}
    \caption{List of the various instruments with their relevant specifications that were used for the analysis of the specific flares.}
    \label{tab:data}
\end{table}
%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations and Data Analysis}\label{sec:obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We study the two flares observed on 28 October, 2021 (X-class) and 29 November, 2020 (M-class) observed on the disk center and limb of the Sun, respectively, from the perspective of Atmospheric Imaging Assembly onboard the {\it Solar Dynamic Observatory} \citep[{\it SDO}/AIA,][]{sdo,aia} vantage point. For this study, we primarily use the observations from AIA, the Extreme Ultraviolet Imager onboard {\it Solar Terrestrial Relations Observatory-A} \citep[{\it STEREO-A}/EUVI,][]{stereo,euvi} and the Solar Ultraviolet Imager onboard {\it GOES} \citep[{\it GOES}/SUVI,][]{suvi}, along with imaging from the Spectrometer/Telescope for Imaging X-rays onboard {\it Solar Orbiter} \citep[{\it SO}/STIX,][]{so,stix,stix1}, the X-Ray Telescope onboard {\it Hinode} \citep[{\it Hinode}/XRT,][]{xrt} and the X-Ray Sensor onboard {\it GOES} \citep[{\it GOES}/XRS,][]{xrs}. For the relevant specification of the instruments used, refer to Table~\ref{tab:data}.

The positions of various spacecraft with respect to the flares' location on the solar disk is illustrated in Fig.~\ref{fig:sc_pos}. The AIA 193 {\AA} observation of the 2021 Oct 28 flare is projected to Heliographic Stonyhurst (HGS) coordinates in panel a, with {\it SDO} (magenta cross), {\it SO} (cyan circle) and {\it STEREO-A} (green triangle) positions marked with respect to the flare location in Fig.~\ref{fig:sc_pos}.a. The flare is marked with white dotted circle. The relative separation between the spacecrafts show the difference in the polar angle (vertical axis in Fig.~\ref{fig:sc_pos}.a) and azimuthal angle (horizontal axis in Fig.~\ref{fig:sc_pos}.a) introduced due to difference in vantages. The positions of the spacecrafts projected onto the Heliocentric Intertial (HCI) frame for the 2021 Oct 28 event is plotted in Fig.~\ref{fig:sc_pos}.b. The relative separation between the spacecrafts show the difference in azimuthal angle and the distance from the sun (distance from the center radially outwards in Fig.~\ref{fig:sc_pos}.b). We show the same for the 2021 Nov 29 flare in Fig.~\ref{fig:sc_pos}.c and d, with {\it STEREO-A}/EUVI 171 {\AA} emission projected to HGS and HCI coordinates.% {\bf We list out the various instruments used for the analysis of the two flares in Tab.~\ref{tab:data}.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X flare on 28th October, 2021}\label{sec:x1-obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%--------------------------
\begin{figure}[ht!]
    %\hspace{-1cm}
    \centering
    \includegraphics[width=0.95\textwidth,trim={0.3cm 4cm 0cm 2cm},clip]{oct28_align_new.pdf}
    \caption{X-class flare observed on October 28, 2021. Panel (a): the post eruption arcade as observed in AIA 193~{\AA}. Panel (b): Two ribbon of the flare observed in AIA 1600 {\AA}, overplotted with STIX soft X-ray contours of 4{--}15 keV (solid blue lines) and hard X-ray contours of 15{--}25 keV (dashed red lines). Panel (c): the emission measure map of the region integrated over the temperature range of 5$<\log\,T<$7.4. Panel (d): the obtained DEM weighted temperature.}
    \label{fig:flare}
\end{figure}
%%--------------------------
The X1 flare occurred on Oct 28, 2021 (SOL2021-10-28T15:35) near $\sim~[0\arcsec,-500\arcsec]$ solar coordinates as seen from AIA. The flare was observed by {\it SDO}/AIA, {\it STEREO-A}/EUVI, {\it GOES}/SUVI, and XRT, along with {\it GOES}/XRS and STIX observations. AIA observations provide imaging with 94, 131, 193, 171, 211 and 335 {\AA} passbands. SUVI provides imaging with similar wavelengths as AIA: 94, 131, 195, 171, 195 and 284 {\AA}, while {\it STEREO-A}/EUVI provides imaging in 171, 195, 284 {\AA} coronal passbands. The flare exhibits a standard two-ribbon structure, along with a ridge at the top of the rising flare arcade \citep{longcope22}. The ridge most prominently shows up in the AIA 193 {\AA} channel with contributions from \ion{Fe}{12} and \ion{Fe}{24}. %\cite{longcope22} showed that the ridge was mainly created by the hotter \ion{Fe}{24} line (T~$\simeq$~17~MK) and a height of the flare arcade, $h~\simeq~20~Mm$. We use this ridge to identify the loop-top in the AIA observation and calculate the height of the loop (see \S\ref{sec:los} for further details).

{\it SO}/STIX provides X-ray imaging and spectroscopy in the 4-150 keV range. Spectroscopic observations are provided with 32 pre-defined energy channels distributed in this energy range. STIX is an indirect imager measuring visibilities in the Fourier space, and images can be reconstructed from these visibilities with different algorithms. We use the MEM-GE algorithm \citep{massa20} to construct the STIX intensity maps.

In Figure.~\ref{fig:flare} (top row) we display the flare as observed with AIA~193 {\AA} (panel a) and 1600~{\AA} (panel b, inverted colors). These images reveal the two ribbon nature of the flare with post eruption arcades \cite[see e.g.][]{TriBC_2004}. We also overplot the STIX intensity contours (15:27:40 {--} 15:28:00 UT) in 4{--}15 keV (blue solid lines) and 15{--}25 keV (red dashed lines), on an AIA 1600 {\AA} image during the rise phase of the flare (see Fig.~\ref{fig:flare}.b). The STIX hard X-ray contour aligns with AIA 1600 {\AA} ribbons.

%%--------------------------
\begin{figure}[ht!]
\centering
    \includegraphics[trim={4.8cm 2.5cm 6cm 2.5cm},clip,width=0.8\textwidth]{nov_29_align.pdf}
    \caption{M-class flare on Nov 29, 2020 as observed nearly simultaneously by AIA~{131 \AA} (panel a) and XRT Be-med filer (panel b) and the derived EM map (panel c) and DEM weighted temperature (panel d). The EM map is obtained by integrating over the temperature range of 5$<\log\,T<$7.4.}
    \label{fig:flare2}
\end{figure}
%%--------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{M flare on 29th November, 2020}\label{sec:m-obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The M flare occurred on November 29, 2020, near $\sim [-950\arcsec,-430\arcsec]$ solar coordinates from the AIA perspective. In the AIA images, the footpoints were occulted behind the limb. The flare was also recorded by {\it STEREO-A}/EUVI, {\it GOES}/SUVI and {\it Hinode}/XRT. Along with these instruments, there were observations from {\it GOES}/XRS and {\it Fermi}/GBM. There were no observations from STIX for this event.

The flare exhibits a fan which is visible in the hot channels of AIA (131 and 193 {\AA}) and the XRT channels. Fig. \ref{fig:flare2}a shows an AIA 131 {\AA} image of the flare recorded during the decay phase. In Fig.~\ref{fig:flare2}b, we plot an XRT Be-med filter observation taken near simultaneous with the AIA image. For the M-class event, the {\it Fermi} observations were periodically eclipsed during the duration of the flare. We do not calculate the cumulative non-thermal energy for this event, as during the eclipses there is no reliable way to estimate the instantaneous non-thermal energy deposited. Hence, the estimation of cumulative non-thermal energy would not be a fair representation of the total non-thermal energy available.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thermal Energy}\label{sec:therm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We estimate the thermal energy of the X-class event by calculating the differential emission measure (DEM) of the flaring region from the six AIA passbands including 94 {\AA} (\ion{Fe}{10} $\sim$1.1~MK, \ion{Fe}{18} $\sim$7.1~MK), 131 {\AA} (\ion{Fe}{8} $\sim$0.4~MK, \ion{Fe}{21} $\sim$11~MK), 171 {\AA} (\ion{Fe}{9} $\sim$0.6~MK), 193 {\AA} (\ion{Fe}{12} $\sim$1.6~MK, \ion{Fe}{24} $\sim$17~MK), 211 {\AA} (\ion{Fe}{14} $\sim$2~MK), 335 {\AA} (\ion{Fe}{16} $\sim$2.5~MK) \citep{o'dwyer10,O'Dwyer12}. Near the peak of the flare, several AIA frames are saturated in several instances and as such they cannot be used for the analysis. The SUVI temperature response is very similar to AIA~\citep{suvi}. Therefore, whenever necessary, the specific AIA frames can be replaced by those taken by SUVI for thermal energy estimates after co-aligning and co-registering  the two observations. Moreover, we also need to rescale the AIA temperature response functions to account for the binning of the AIA data to match the SUVI resolution. Note that since most of the XRT observations during the flare are saturated, we do not include them in our analysis of this flare.

For the M-class event, we estimate the thermal energy using AIA and XRT observations. Here, we bin the AIA observations to XRT pixel size, co-align, and calculate the DEMs. As described earlier, we rescale the AIA temperature response functions to account for the binning of the AIA data.

We use the regularized inversion method \citep{hannah&kontar12} to calculate the DEMs. The observed flux from the individual passbands can be written as,
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
    F_{i}~=~\int~DEM_{c}(T)~R_{i}(T)~dT+\delta F_{i}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%
where $R_{i}$ is the temperature response of the passband `i', $\delta F_{i}$ is the error on the observed intensity for the passband `i' and $DEM_{c}(T)$ is the column differential emission measure (in units of $cm^{-5}K^{-1}$). The DEM inversion was performed for a temperature range $4.5~\le log(T)\le~8$.

With the inverted DEM, we calculate the column emission measure ($EM^{c}~\mathrm{in~units~of~cm^{-5}}$) and DEM-weighted temperature as:
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
    EM^{c}~=~\int_{T} DEM_{c}(T)~dT
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
    \bar{T}~=~\frac{1}{EM^{c}}\int_{T} DEM_{c}(T)~T~dT
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%

%%--------------
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth,trim={2cm 6cm 2cm 6cm},clip]{flare_orient.pdf}
    \caption{Panel (a): {\it STEREO-A}/EUVI 171 {\AA} observation of the flare arcade in the decay phase. The magenta cross marks a point on the top of the arcade. The line of sight (LOS) goes into the page through that point. Panel (b): SDO/AIA 171 {\AA} observation of the flare arcade. The red line marks the LOS through the arcade from the STEREO-A perspective projected to the AIA perspective.} %Panel (c): Cartoon demonstrating the projection effects of a semicircular loop between STEREO-A and AIA vantages due to the difference between polar ($\theta$) and azimuthal ($\phi$) angles respectively. Panel (d) : The calculated LOS map from the AIA perspective, with the calculated loop top height under the assumption of a semi-circular loop geometry.}
    \label{fig:flare_orient}
\end{figure}
%%--------------
%%--------------
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\textwidth,trim={4cm 5cm 3cm 5cm},clip]{nov29_triangulate.pdf}
    \caption{Panel (a): SDO/AIA 193 {\AA} observation of the Nov 29 flare in the decay phase. The blue cross marks a point on the top of the arcade. The LOS goes into the page from AIA perspective. Panel (b): {\it STEREO-A}/EUVI 195 {\AA} observation around same time. The blue solid line is the LOS from AIA perspective in panel (a) projected onto {\it STEREO-A} perspective. We can trace out the height of the flare arcade at various locations using \textit{scc\_measure.pro}. The height of the arcade at various points from the Sun's surface is marked in panel (a).}
    \label{fig:flare_orient2}
    \end{figure}
%%--------------

The temperature range of the integration is set to $5 < \log\,T < 7.4$. We calculate and plot the emission measure (panel c in Figs.~\ref{fig:flare} \& \ref{fig:flare2}) and the DEM weighted temperature (panel d in Figs.~\ref{fig:flare} \& \ref{fig:flare2}) for all the pixels in the field of view (FOV) for both the flares.

The main goal of calculating the DEM maps is to estimate the thermal energy (using Eqn.~\ref{eq:t_eneg_1}) arising from various parts of the flare arcade. As alluded to earlier, one of the most crucial quantity is to obtain the volume of the flaring plasm. For our analysis, the volume of the plasma along an individual pixel is given by $V_{j}~=~A\times~LOS_{j}$, where $LOS_{j}$ is the line of sight (LOS) for the j-th pixel in the FOV and $A$ is the physical area of a pixel.

%%--------------
\begin{figure}[ht!][!h]
    \centering
    \includegraphics[width=0.8\textwidth,trim={3cm 2cm 3cm 0cm},clip]{flare_orient_pt2.pdf}
    \caption{Panel (a): Cartoon demonstrating the projection effects of a semicircular loop between STEREO-A and AIA vantages due to the difference between polar ($\theta$) and azimuthal ($\phi$) angles, respectively. Panel (b) : The calculated LOS map from the AIA perspective for the 28th October, 2021 event, with the estimated loop top height under the assumption of a semi-circular loop geometry.}
    \label{fig:flare_orient_2}
\end{figure}
%%--------------

We use the above expression for volume along with the emission measure and DEM weighted temperature estimated from the inverted DEM in Eqn.~\ref{eq:t_eneg_1} and sum over the pixels in the FOV to estimate the thermal energy arising from the FOV. This procedure gives us,
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
    U_{Th}~=~\sum_{j}~\frac{3k_{B}A\sqrt{LOS_{j}}}{\sqrt{EM^{c}_{j}}}\int_{T}~DEM_{c}(T)_{j}T~dT
    \label{eq:t_eneg}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $EM^{c}_{j}$ and $DEM_{c}(T)_{j}$ are the column emission measure and the column DEM in 5.0$<$$\log\,T$$<$7.4 for the j-th pixel in the FOV. We have used $n_{e}^{2}\times LOS~=EM^{c}$ and a filling factor $\textit{f}~=~1$. For the 2020 Nov 29 flare we mask the solar disk while calculating the DEMs and thermal energy. As evident from Eqn.~\ref{eq:t_eneg}, one of the major uncertainties in determining the emitting volume depends on determining the LOS for individual pixels in the FOV.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Determining the LOS}\label{sec:los}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To infer the LOS along individual pixels in the FOV for both the flares, we use AIA and STEREO-A/EUI observations, recorded from two different vantage points, to calculate the height of the flare loop. We display STEREO-A/EUI and AIA-171 image taken neeary simultaneously in Fig.~\ref{fig:flare_orient} a~\&~b, respectively. The magenta cross on Fig.~\ref{fig:flare_orient}.a shows a point at the top of the flare arcade. The LOS goes into the page through that point from the {\it STEREO-A} vantage. The red line in Fig.~\ref{fig:flare_orient}.b shows the LOS through the point in Fig.~\ref{fig:flare_orient} a projected to {\it SDO}/AIA point of view. This demonstrate the geometric effect of observing from different vantage points.

We show the same in Fig.~\ref{fig:flare_orient2} for the Nov 29, 2020 flare. The blue cross marked in Fig.~\ref{fig:flare_orient2}.a is the LOS going into the page from AIA perspective. The blue line marked in Fig.~\ref{fig:flare_orient2}.b is the same LOS projected onto the STEREO-A perspective. Distances on the observation from two different vantages would be projected by the differences in the polar and the azimuthal angle between the two vantages. Fig.~\ref{fig:flare_orient_2}.a shows the projection effect of a semicircular loop between AIA and STEREO-A LOS. Depending on the orientation of the loops, the projection along the polar and azimuthal angle projects the length of the loop and the height of the loop. The height at various position of the flare arcade is marked in Fig.~\ref{fig:flare_orient2}.a. We assume a semi-circular loop geometry with the height of the loop top to calculate the LOS along every pixel within the flare arcade.

We use co-temporal {\it STEREO-A} 171 {\AA}, 195 {\AA}, and AIA~171 {\AA}, 193 {\AA} images with \textit{scc\_measure.pro}, available in the \textit{sswidl}, to calculate the height of the loop. This routine allows us to select a point on the observation from one of the two vantages. The line of sight (LOS) through the same point is shown on the observation from the different vantage. The same point can be identified on the projected LOS by the user, by identifying similar emission characteristics. The program determines the 3D coordinates of the point (heliographic latitude, longitude and radial distance). We carry out the same measurement at various positions along the loop top, to calculate the change in loop height across the arcade. With the different height measurements at various locations, we can calculate the volume of the flaring plasma under a semicircular assumption. We note that the height estimation using \textit{scc\_measure.pro} is limited by our identification of ``similar emission characteristics" between the two observations from AIA and {\it STEREO-A}.

The extent of the flare arcade is calculated by selecting pixels from the FOV which are located within an emission measure contour of 5\% of the peak emission measure value. It is possible to calculate the extent of the flare arcade from the intensity observed from any of the {\it SDO}/AIA or STIX imaging (for further details about the STIX imaging procedure, please refer to \cite{massa20}). However, due to the highly multi-thermal nature of the plasma, this procedure might be prone to missing significant portions of the flare arcade. The emission measure inferred from the DEM, on the other hand, reflects the density of the plasma in the entirety of the concerned temperature range of $5<\log\,(T)<7.4$. Therefore we find that, a 5\% contour of peak emission measure gives a better estimate of the extent of flare arcade. Outside of the flare arcade, we assume a LOS~$\sim$~2 Mm (the average thickness of chromosphere).

Figure \ref{fig:flare_orient_2}.b shows an example of the calculated LOS in the FOV for the 2021 October 28 flare. The highest point of the flare arcade is oriented from south east towards north west, following the ridge. We get an estimate of the highest point of the loop $\simeq$ 22~Mm, which is consistent with the estimates of the height, $H~\simeq~20~Mm$ of the ridge from magnetic loop modelling by \cite{longcope22} for the same event. Using the map of the LOS shown in the bottom panel of Fig. \ref{fig:flare_orient_2}, we determine the thermal energy in every pixel of the flare using Eqn.~\ref{eq:t_eneg}.  A similar process is used for the flare loops in the 2020 November 29 flare.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy in the non-thermal electrons}\label{sec:non-therm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We estimate the energy in the non-thermal electrons for the 2021 Oct 28 X-class event by fitting the STIX spectra. In Fig.~\ref{fig:stix_an}.a, we plot the STIX light curve of the event in 4 to 6 keV (solid black line), 6 to 12 keV (solid magenta line), 12 to 25 keV (solid yellow line) and 25 to 50 keV (solid blue line). The attenuator kicked in around $\sim$ 15:28 UT, which drastically cuts down the intensity to avoid saturation. The hard X-ray peak is visible in the 25 to 50 keV (solid blue line) at $\sim$ 15:28 UT. The soft X-ray peak is visible within the attenuated flux in 6 to 12 keV (solid magenta line) at $\sim$ 15:30 UT.
